---
title: "PSYC7005 Week 4"
author: "Lucy Harlow"
date: "12/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A Statistics Project for PSYC7005

I decided to supplement my learning of jamovi by attempting the worksheet exercises in R as well. My main source for learning R continues to be swirl, but working on these exercises will support this learning with practical experiences. It took quite a while to figure out how to use ggplots, and I didn't quite get the boxplot to show what I wanted---depicting the points proved challenging, so I'll come back to this later when I understand things more fully.


```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
winningmargins <- read.csv("aflsmall_margins.csv", header = TRUE)
hist(winningmargins$afl.margins, main = "Team Winning Margins", 
     xlab = "Team Winning Margins", breaks = 20, freq = FALSE)
ggplot() +
  geom_boxplot(data = winningmargins,
               aes(x = afl.margins, y = ""),
  ) + labs(x = "Team Winning Margins", y = "") + coord_flip() +
  geom_point(mapping = NULL)

```

